
## Packages
```{r packages}
library(tidyverse)
```

## Load the data
```{r data}

mydata_c <-  read_csv("Data/control_latest.csv")
mydata_e <-  read_csv("Data/events_latest.csv")

```

## Line Chart
Variables needed:

- `date`
- count for events
- initiated


- two vertical lines: start date in range, end date in range

- horizontal time axis from the first to the last day in the dataset

- vertical axis being the number of events

- two separate lines (shapes and colors): one for events initiated by the  Ukrainian forces, the other for events initiated by the Russian forces

Steps:
1: filter out (remove) events initiated by neither party
2: filter out other events
3: count the number of events

```{r line data prep}
mydata_line_plot <- mydata_e

# example input
year_start <- 20220224
year_end <- 20220312
event_selected <- c("t_airstrike_b", "t_armor_b")

# prepare a binary variable indicating which of the two sides initiated the event
mydata_line_plot$initiated <- case_when(
  mydata_line_plot$a_rus_b == 1 ~ "Russia-initiated",
  mydata_line_plot$a_ukr_b == 1 ~ "Ukraine-initiated",
)

# filter out (remove) events initiated by neither party
mydata_line_plot <- mydata_line_plot %>% filter(is.na(mydata_line_plot$initiated) == FALSE)

# to subset the data to the selected variables:
# we generalize this:
# mydata_e <- mydata_e %>% subset(t_airstrike_b!=0 | t_armor_b!=0) 
# to this:
# mydata_line_plot <- mydata_line_plot %>% subset(get(event_selected[1])!=0 | get(event_selected[2])!=0)
# thanks to a very useful get() command: https://cran.r-project.org/doc/FAQ/R-FAQ.html#How-can-I-turn-a-string-into-a-variable_003f
# we further generalize for any number of event variables selected by the user: 
### any_of() takes a character vector with the names of the columns we want to consider. Then if_any() takes the columns as arguments to check if at least one of them meets the specified condition.
mydata_line_plot <- mydata_line_plot %>% filter(if_any(any_of(event_selected)) == 1)

# count the number of events
mydata_line_plot <- mydata_line_plot %>% count(date, initiated, sort = TRUE)

# calculate the cumulative sum over time, grouped by the country that initiated the event
# https://datacornering.com/cumulative-sum-or-count-in-r/
mydata_line_plot <- mydata_line_plot %>%
  group_by(initiated) %>%
  arrange(date, decreasing = FALSE) %>%
  mutate("n_cum" = cumsum(n))
```

```{r line}
ggplot(mydata_line_plot, aes(x = date, y = n_cum, color = initiated, linetype = initiated)) +
geom_line() +
geom_vline(xintercept = year_start, color = "grey50") +
geom_vline(xintercept = year_end, color = "grey50")
```

## Bar chart
Assign events to specific regions with the help of spatial joint and a shapefile of Ukraine divided into regions.
horizontal bars
Regions of ukraine on the left axis
Total number of all selected event types for the specified time range
Each bar divided into different colors corresponding to different event types

```{r bar}
# filter out events outside of the date range
# IS THIS CORRECT? SIMPLY PUTTING GREATER OR EQUAL?
# mydata_e <- mydata_e %>% filter(date <= year_end & date >= year_start)
```