---
title: "Project 2 Proposal"
author: "Group Name: other-carp"
output: github_document
---
```{r load-packages, echo = FALSE, include = FALSE, message=FALSE}
library(tidyverse)
library(styler)
library(here)
```

# Introduction

Russian-Ukraine crisis is a long lasting problem and it caused countless casualties nearly everyday. The earliest big confrontation can be dated back to 2014, when two cities from Ukraine (Donetsk and Luhansk) started revolution. From 2014 to 2022, there are several small military events and non-military but violent events happened at the boarder of Russia-Ukraine. In 2022, concerning the East expansion of NATO, Russia started a war against Ukraine and brings up the confrontation. Many people are killed and many cities are destroyed during the war. In this project, we want to build a shiny app to illustrate the military events happened during the recent 2022 Russia-Ukraine war and keep track of the territory boundary or military marching of soldiers from two sides.

We choose this topic because all of us are concerned about international confrontations. We hope that the world is peaceful and there is no war or violent event exist any more. Also, it would be interesting to take a look at the which place has the most attackers or military actions, and which side has the relative advantages as the battles continued. If we can plot every single events on a map, we will be able to discover some patterns such as which area or city has the most violent event happen and hopefully, we can prevent the same attacks from happening again. 

What we want to do is to use shiny app and interactive plots to live demonstrate the process of Russia-Ukraine war, including the territory movement, military and non-military events across different time frame. The GOAL is to create pieces of "generative" art, ie., pieces of art that have been generated using computer input and mathematical formulas. You can find more details on data frame below and all the codes are included on "app.R" file in Shiny App folder.


# Data
We will use the VIINA (Violent Incident Information from News Articles on the 2022 Russian Invasion of Ukraine) data set collected by Professor Yuri Zhukov from the University of Michigan (see [full list here](https://github.com/zhukovyuri/VIINA)). This is the most comprehensive data set from the internet, which provides us extensive information to visualize, and to recover deep insights.

It's a near-real time multi-source event data system based on news reports from Ukrainian and Russian media, and Zhukov geocoded and classified them into  standard conflict event categories through machine learning. The events are GIS-ready, with temporal precision down to the minute. The final data set was orginized and saved into two csv files.

## Event data

```{r}
event <- read.csv(here("Data", "events_latest.csv"))
head(event)
```

The first csv file includes raw events, and below is the codebook:

### Codebook
- event_id: Unique event ID
- report_id: Unique ID for report that contains the event
- location: Index of unique locations mentioned in each event
- tempid: Temporary numeric ID
- source: Data source short name
- date: Date of event report (YYYYMMDD)
- time: Time of event report (HH:MM)
- url: URL web address of event report
- text: Text of event report headline/description
- lang: Language of report (ua is Ukrainian, ru is Russian)
- address: Address of geocoded location
- longitude: Longitude coordinate of event location
- latitude: Latitude coordinate of event location
- GEO_PRECISION: geographic precision of geocoded location
- GEO_API: Geocoding API used to locate event
- t_[event type]: Predicted probability that report describes event of each type (from LSTM model, see above)
- a_[actor]: Predicted probability that report describes event initiated by each actor (from LSTM model, see above)

The data includes following event categories:

### Categories
- t_mil: Event is about war/military operations
- t_nmil: Event is not about war/military operations (e.g. human interest story)
- t_loc: Event report includes reference to specific location
- t_san: Event report mentions economic sanctions imposed on Russia
- a_rus: Event initiated by Russian or Russian-aligned armed forces
- a_ukr: Event initiated by Ukrainian or Ukrainian-aligned armed forces
- a_civ: Event initiated by civilians
- a_other: Event initiated by a third party (e.g. U.S., EU, Red Cross)
- t_aad: Anti-air defense, Buk, shoulder-fired missiles (Igla, Strela, Stinger)
- t_airstrike: Air strike, strategic bombing, helicopter strike
- t_armor: Tank battle or assault
- t_arrest: Arrest by security services or detention of prisoners of war
- t_artillery: Shelling by field artillery, howitzer, mortar, or rockets like Grad/BM-21, Uragan/BM-27, other Multiple Launch Rocket System (MRLS)
- t_control: Establishment/claim of territorial control over population center
- t_firefight: Any exchange of gunfire with handguns, semi-automatic rifles, automatic rifles, machine guns, rocket-propelled grenades (RPGs)
- t_ied: Improvised explosive device, roadside bomb, landmine, car bomb, explosion 
- t_raid: Assault/attack by paratroopers or special forces, usually followed by a retreat
- t_occupy: Occupation of territory or building
- t_property: Destruction of property or infrastructure
- t_cyber: Cyber operations, including DDOS attacks, website defacement
- t_hospitals: Attacks on hospitals and humanitarian convoys
- t_milcas: Event report mentions military casualties
- t_civcas: Event report mentions civilian casualties

## Territorial control
```{r}
control <- read.csv(here("Data", "control_latest.csv")) %>%
  select(1:8)
head(control)
```

Based on the event data set, Zhukov collected a csv file indicating whether each district administrative center is presently under the control of Ukrainian forces, Russian forces, or is being actively contested between the two. This is on day - place level.

Below is the code book:

### Codebook

- geonameid: Numeric ID of populated place
- name: Name of populated place
- asciiname: Name of populated place, ASCII values
- alternatenames: Alternative spellings of place name
- longitude: Longitude coordinate of populated place
- latitude: Latitude coordinate of populated place          
- feature_code: Type of populated place (see [full list here](https://www.geonames.org/export/codes.html))
- ctr_[YYYYMMDDHHMMSS]: Control status, with timestamp (UA/RU/CONTESTED)


# Analysis/Graph Outline
## The more specific goals of our visualization and analysis
In our visualization, we would like to combine the information from both datasets to create one map with switchable layers. As a result, the user will be able to see both the events and the territorial division at a given point in time (rather than separately like in the dataset author's own visualizations https://github.com/zhukovyuri/VIINA). Moreover, our goal is to vastly improve the number and quality of features as well as aesthetics as compared to the existing Shiny App based on the VIINA project: https://schardata.shinyapps.io/viina/. We hope that the end result of our project will be attractive to the author of the dataset, his original audience, and broader groups of people interested in the past and current states of the ongoing war.

In our analysis, we are especially interested in two broad questions:

1. How do the sides of the conflict compare? In particular, how they have been affected relatively to each other in terms of the type and the number of events; how does their activity differ, especially in the case of events that do not depend on being the invader vs. the the defender; how does the control by a given side impact the event type occurrence?

2. How ...

	• Which regions were most affected?
		○ event types
		○ event count
		○ event density
	• How has the war been developing? Spatial network analysis.
		○ directionality in events by type over time
		○ spatial or temporal relationships between event types
			• distance from point X
	• How will the war develop?
		○ (simple model of) extrapolation of trends from X days and in X regions

## Tools of analysis
	• network analysis
		○ centrality
		○ connectedness
		
		
If the events are nodes are ... and ties are ..., then .... .  


	• Count the time controlled by each party and visually represent it
		○ perhaps as alpha or a scale that goes from one color to the next

## Graph details

Ideas for graph(s):
	• timeline interval width selection and shifting
	• show the time when X percent was under the Russian occupation
	• show the time when region X was under the Russian occupation
	• Event counter to date or between dates
	• Indicate/filter source origin (at least case by case, like in the original viz)
	
Graphs generated as shots of what the interactive viz shows

The graph will show the map of Ukraine as a part of a larger world atlas (like OpenStreetMap). On the map, there will be a timeline axis with a scalable and shiftable interval. The endpoints of the interval will be able to be determined numerically as well.

Below the graph, there will be feature selectors. There will be the following primary features:
	• show the territorial division (based on the territorial control dataset). 
		○ show or highlight only one of the sides of conflict

And we will add some of the following additional features:
	• save the map as an image file
	• show event details
		○ title
		○ type of event
		○ text
		○ time
		○ address
		○ lon/lat of event
		○ url
	• show only event reports of types X, Y, Z
		○ a tree: military/non-military boxes and then each of the more specific category boxes
		○ boxes independent of military/non-military classification (e.g. by whom it was initiated)

	• not just a map, but also:
		○ barplot()
		○ graph of events over time and facet by location?
		○ or geofacet()?

In terms of geoms:
	• geom_point(), of course
	• geom_label()?
	• geom_density_2d() with mods
	• choropleth AND heatmap as derivatives of maps with points (e.g. counts)


# Limitations
How ... The location or affiliation of the news source ... reported may be crucial for understanding the results ... 
		○ add source origin (Ukrainian vs. Russian, regardless of language)
		enhancing the dataset itself is beyond the scope of our project.

# Schedule

## Week 2 (May 9 - 15)  
- write proposal
- organize data and generate variables
  
## Week 3 (May 16 - 22)  
- plot maps

## Week 4 (May 23 - 29)  
- make shiny app
- write-up

## Week 5 (May 30 - )  
- present

# Github Organization
We have three folders: Data, Proposal and Shiny App. You will find our latest data sets regarding the events and controlled area under Data folder. You will find our proposal under Proposal folder. You will find our demonstration under Shiny App folder.


# References
Zhukov, Yuri (2022). "VIINA: Violent Incident Information from News Articles on the 2022 Russian Invasion of Ukraine." Ann Arbor: University of Michigan, Center for Political Studies. (https://github.com/zhukovyuri/VIINA, accessed 05/11/2022).



