---
title: "Project 2 Proposal"
author: "Group Name: other-carp"
output: github_document
---
```{r load-packages, echo = FALSE, include = FALSE, message=FALSE}
library(tidyverse)
library(styler)
library(here)
```

# Introduction

In this project, we build a Shiny app with interactive plots to visualize Russian-Ukraine crisis, which help calling people's attention, and recover deep insights to make peace.

Russian-Ukraine crisis is a long lasting problem and it caused countless casualties nearly everyday. The earliest big confrontation can be dated back to 2014, when pro-Russian separatists in two cities in Ukraine, Donetsk and Luhansk, started revolts. From 2014 to 2022, there were several small military events and non-military but violent events happened in Ukraine in the area close to the Russian-Ukrainian border. In 2022, allegedly concerned with the potential Eastward expansion of NATO, Russia started a war against Ukraine by attacking targets in the whole country using air weapons and by invading its land territory from multiple sides at the same time. Millions of refugees fled the country, dozens of thousands of people including civilians have been killed in targeted and untargeted attacks, many people have been tortured and robbed of their life assets, and many cities have been destroyed within several weeks. In this project, we want to build a shiny app to illustrate the military events happened during the recent 2022 Russia-Ukraine war and keep track of the territory boundary or military marching of soldiers from two sides.

We have chosen this topic, because all of us are concerned about international confrontations. We wish the world were peaceful and there were no wars or violent events anymore. 

We are also motivated by the opportunity of enhancing the study of the war in its specific aspects, such as place has the most attackers or military actions, and which side has the relative advantages as the battles continue. If we can plot every single events on a map, we will be able to discover some patterns such as which area or city has the most violent event happen. Perhaps insights with tools such as ours could be somehow used prevent the same attacks from happening again. 

We used interactive plots to live demonstrate the process of Russia-Ukraine war, including the territory movement, military and non-military events across different timeframes. 

Here is the link to our shiny app:
https://wdxbh1998.shinyapps.io/UkrainWar

# Data
We used the VIINA (Violent Incident Information from News Articles on the 2022 Russian Invasion of Ukraine) data set collected by Professor Yuri Zhukov from the University of Michigan (see [full list here](https://github.com/zhukovyuri/VIINA)). This is the most comprehensive data set from the internet, which provides us extensive information to visualize.

It's a near-real time multi-source event data system based on news reports from Ukrainian and Russian media, and Zhukov geocoded and classified them into  standard conflict event categories through machine learning. The events are GIS-ready, with temporal precision down to the minute. The final data set was orginized and saved into two csv files.

## Event data

```{r}
event <- read.csv(here("Data", "events_latest.csv")) %>%
  select(6:18) %>%
  select(!address)
head(event)
```

The first csv file includes raw events, and below is the codebook:

### Codebook
- event_id: Unique event ID
- report_id: Unique ID for report that contains the event
- location: Index of unique locations mentioned in each event
- tempid: Temporary numeric ID
- source: Data source short name
- date: Date of event report (YYYYMMDD)
- time: Time of event report (HH:MM)
- url: URL web address of event report
- text: Text of event report headline/description
- lang: Language of report (ua is Ukrainian, ru is Russian)
- address: Address of geocoded location
- longitude: Longitude coordinate of event location
- latitude: Latitude coordinate of event location
- GEO_PRECISION: geographic precision of geocoded location
- GEO_API: Geocoding API used to locate event
- t_[event type]: Predicted probability that report describes event of each type (from LSTM model, see above)
- a_[actor]: Predicted probability that report describes event initiated by each actor (from LSTM model, see above)

The data includes following event categories:

### Categories
- t_mil: Event is about war/military operations
- t_nmil: Event is not about war/military operations (e.g. human interest story)
- t_loc: Event report includes reference to specific location
- t_san: Event report mentions economic sanctions imposed on Russia
- a_rus: Event initiated by Russian or Russian-aligned armed forces
- a_ukr: Event initiated by Ukrainian or Ukrainian-aligned armed forces
- a_civ: Event initiated by civilians
- a_other: Event initiated by a third party (e.g. U.S., EU, Red Cross)
- t_aad: Anti-air defense, Buk, shoulder-fired missiles (Igla, Strela, Stinger)
- t_airstrike: Air strike, strategic bombing, helicopter strike
- t_armor: Tank battle or assault
- t_arrest: Arrest by security services or detention of prisoners of war
- t_artillery: Shelling by field artillery, howitzer, mortar, or rockets like Grad/BM-21, Uragan/BM-27, other Multiple Launch Rocket System (MRLS)
- t_control: Establishment/claim of territorial control over population center
- t_firefight: Any exchange of gunfire with handguns, semi-automatic rifles, automatic rifles, machine guns, rocket-propelled grenades (RPGs)
- t_ied: Improvised explosive device, roadside bomb, landmine, car bomb, explosion 
- t_raid: Assault/attack by paratroopers or special forces, usually followed by a retreat
- t_occupy: Occupation of territory or building
- t_property: Destruction of property or infrastructure
- t_cyber: Cyber operations, including DDOS attacks, website defacement
- t_hospitals: Attacks on hospitals and humanitarian convoys
- t_milcas: Event report mentions military casualties
- t_civcas: Event report mentions civilian casualties

## Territorial control
```{r}
control <- read.csv(here("Data", "control_latest.csv")) %>%
  select(1:8)
head(control)
```

Based on the event data set, Zhukov collected a csv file indicating whether each district administrative center is presently under the control of Ukrainian forces, Russian forces, or is being actively contested between the two. This is on day - place level.

Below is the code book:

### Codebook

- geonameid: Numeric ID of populated place
- name: Name of populated place
- asciiname: Name of populated place, ASCII values
- alternatenames: Alternative spellings of place name
- longitude: Longitude coordinate of populated place
- latitude: Latitude coordinate of populated place          
- feature_code: Type of populated place (see [full list here](https://www.geonames.org/export/codes.html))
- ctr_[YYYYMMDDHHMMSS]: Control status, with timestamp (UA/RU/CONTESTED)


# Analysis Outline
## The goals of our analysis

In our analysis, we are especially interested in two broad questions:

1. How do the sides of the conflict compare? In particular, how they have been affected relatively to each other in terms of the type and the number of events; how does their activity differ, especially in the case of events that do not depend on being the invader vs. the the defender(what we mean here is that along with the war, not only Russian-Ukraine troops harm civilians, but other non-military events such as robbery); how does the control by a given side impact the event type occurrence?

2. How can the events improve our spatial understanding of the war? More specifically, how do areas of Ukraine differ in terms of event type, count, and density? Which regions were most affected? How have the centers of gravity of the war events (broadly or narrowly defined) shifted and changed their importance over time?

## Minor Remarks Regarding the Tools of Analysis

For the first question, it will be useful to determine the impact each side of the conflict had on a given territory with relation to the events there. This can be achieved by counting the time a region is controlled by each side. Then, on the map, this could be represented by an alpha value or a scale that goes from one color to the next.
 
For the second question, a network analysis can be useful: If the events are nodes and the temporal or spatial distances between them are ties (with strenghts proportional to the distances), then we can obtain the centrality and connectedness of various locations in Ukraine as measures of their importance in the war. These two values, in turn, can be represented on the map as the sizes of the nodes. 

# Visualization Outline

## Map
In our visualization, we would like to combine the information from both datasets to create one map with switchable layers. As a result, the user will be able to see both the events and the territorial division at a given point in time (rather than separately like in the dataset author's own visualizations https://github.com/zhukovyuri/VIINA). Moreover, our goal is to vastly improve the number and quality of features as well as aesthetics as compared to the existing Shiny App based on the VIINA project: https://schardata.shinyapps.io/viina/. We hope that the end result of our project will be attractive to the author of the dataset, his original audience, and broader groups of people interested in the past and current states of the ongoing war.

Graphs will (hopefully) be able to be generated as shots of what the interactive visualization is showing with the specified parameters.

The graph will show the map of Ukraine as a part of a larger world atlas (like OpenStreetMap).

### Map details
In terms of geoms, the map will containt `geom_point()` and `geom_density_2d()` with appropriate modifications. The `geom_label()` function does not seem too useful, but we will consider adding it.

We would like the app to generate maps with both choropleths and heatmaps as derivatives (sort of) of maps with points (e.g. a map with points being a more fundamental version of a density map). 

We are considering the following main features of the map generator:

	• Select and shift of the time interval width on a timeline. On the map, there will be a timeline axis with a scalable and shiftable interval. The endpoints of the interval will be able to be determined numerically as well.
	
	• Indicate and/or filter source origin (at least case by case, like in the original visualization)
	
	• Show the time when X percent was under the Russian occupation
	
	• Show the times when region X was under the Russian occupation on the timeline
	
	• Include an event counter up to a specific date or between dates
	
	• Show the territorial division (based on the territorial control dataset). Show or highlight only one of the sides of conflict

We are also considering the following additional features:

	• save the map as an image file
	
	• show event details
		○ title
		○ type of event
		○ text
		○ time
		○ address
		○ lon/lat of event
		○ url
		
	• show only event reports of types X, Y, Z
		○ Add a tree: military/non-military selector boxes first and then each of the more specific category selector boxes
		○ Add also boxes independent of military/non-military classification (e.g. by whom it was initiated)

## Bar charts

Bar charts will be helpful to generate whenever the user is less interested in the geographic location and more interested in direct or precise comparisons between spatial units. This seems especially relevant for question 1, where the location matters less than a the mere assignment to a particular side of the conflict.

In terms of geoms, we will consider using `geom_bar` and `geom_col()`. We will allow the user to facet by location with `facet_wrap()` function and/or the `geofacet` package.

## Line trends
We will allow the user to generate plots of events over time with `geom_line()` and apply filters to modify narrow the data sample or make cross-group comparisons, such as the number of tank battles or assults by the side controlling the territory in the first month of the conflict.

# Limitations and Suggestions for Future Research
The location or the affiliation of the news source that reported an event may be crucial for understanding the results of analysis, as the freedom of press and the journalistic standards likely vary drastically between Ukraine and Russia. Thus, future research could add a variable and a corresponding app feature that enables the distiction between the Ukrainian and Russian sources. However, enhancing the dataset itself is beyond the scope of our project.

Future research could also develop and implement in a visualized form predictions about the spatial developments of the war over time.

# Schedule

## Week 2 (May 9 - 15)  
- write proposal
- organize data and generate variables
  
## Week 3 (May 16 - 22)  
- plot maps

## Week 4 (May 23 - 29)  
- make shiny app
- write-up

## Week 5 (May 30 - )  
- present

# Github Organization
We have three folders: Data, Proposal and Shiny App. You will find our latest data sets regarding the events and controlled area under Data folder. You will find our proposal under Proposal folder. You will find our demonstration under Shiny App folder.


# References
Zhukov, Yuri (2022). "VIINA: Violent Incident Information from News Articles on the 2022 Russian Invasion of Ukraine." Ann Arbor: University of Michigan, Center for Political Studies. (https://github.com/zhukovyuri/VIINA, accessed 05/11/2022).



