
## Packages
```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 12))
```

## Load the data
```{r data, message=FALSE, warning=FALSE}
mydata_e <-  read_csv("Data/events_latest.csv")
```

## Clean the data
```{r clean}
mydata_e <- mydata_e %>% select(-ends_with("_pred"), report_id, location, tempid, source, url, time, text, lang, address, GEO_PRECISION, GEO_API)
```

## Line Chart

Specification:

- two vertical lines: start date in range, end date in range

- horizontal time axis from the first to the last day in the dataset

- vertical axis being the number of events

- two separate lines (shapes and colors): one for events initiated by the  Ukrainian forces, the other for events initiated by the Russian forces

Variables needed:

- `date`
- count for events
- initiated

Steps:
1: filter out (remove) events initiated by neither party
2: filter out other events
3: count the number of events

```{r line data prep}
mydata_line_plot <- mydata_e

# example input
year_start <- 20220224
year_end <- 20220312
event_selected <- c("t_airstrike_b", "t_armor_b")

# prepare a binary variable indicating which of the two sides initiated the event
mydata_line_plot$initiated <- case_when(
  mydata_line_plot$a_rus_b == 1 ~ "Russia-initiated",
  mydata_line_plot$a_ukr_b == 1 ~ "Ukraine-initiated",
)

# filter out (remove) events initiated by neither party
mydata_line_plot <- mydata_line_plot %>%
  filter(is.na(mydata_line_plot$initiated) == FALSE) %>%
  filter(if_any(any_of(event_selected)) == 1) %>% # count the number of events
  count(date, initiated, sort = TRUE) %>% # calculate the cumulative sum over time, grouped by the country that initiated the event
# https://datacornering.com/cumulative-sum-or-count-in-r/
  group_by(initiated) %>%
  arrange(date, decreasing = FALSE) %>%
  mutate("n_cum" = cumsum(n))
```

```{r line graph}
ggplot(mydata_line_plot, aes(x = date, y = n_cum, color = initiated, linetype = initiated)) +
geom_line() +
geom_vline(xintercept = year_start, color = "grey50") +
geom_vline(xintercept = year_end, color = "grey50")
```

## Bar chart
- Assign events to specific regions with the help of spatial joint and a shapefile of Ukraine divided into regions;
- Horizontal bars;
- Regions of Ukraine on the left axis;
- Total number of all selected event types for the specified time range;
- Each bar divided into different colors corresponding to different event types

```{r data prep}
mydata_col_plot <- mydata_e
shp <- st_read("Data/shp_city/pp624tm0074.shp")

# Assign events to specific regions with the help of spatial joint and a shapefile of Ukraine divided into regions.
mydata_col_plot <- mydata_col_plot %>%
st_as_sf(coords = c("longitude", "latitude"),
           crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>% 
  st_join(shp, left = FALSE) %>%
  st_drop_geometry() %>%
  rename(region = name_1)
```


```{r get and transform user input}
# example input
year_start <- 20220224
year_end <- 20220312
event_selected <- c("t_airstrike_b", "t_armor_b")

mydata_col_plot <- mydata_col_plot %>% filter(if_any(any_of(event_selected)) == 1) %>% # filter out events outside of the date range
  filter(date <= year_end & date >= year_start)

# Count the number of events by event type and region. The following loops over all event types selected by the user to calculate the sum of the events of each type in the filtered data frame.
regions_list <- unique(mydata_col_plot$region)
sums_all_events <- tibble()
for (i in 1:length(regions_list)) {
  temp <- tibble(.rows = 1)
  temp$event_region <- ""
  temp$event_type <- ""
  temp$event_sum <- NA
  for (j in 1:length(event_selected)) {
    temp$event_region <- regions_list[i]
    temp$event_type <- event_selected[j]
    temp$event_sum <- mydata_col_plot %>% 
      filter(region == regions_list[i]) %>%
      select(contains(event_selected[j])) %>% 
      sum()
    sums_all_events <- bind_rows(sums_all_events, temp)
  }
}

mydata_col_plot <- sums_all_events
```

```{r generate the bar chart (column plot)}
# horizontal bars
ggplot(mydata_col_plot, aes(x = event_sum, y = event_region, fill = event_type)) +
  geom_col() +
  labs(y = "Region", x = "Event Type") +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
  )
```